Data Specialization for a Staged Lambda Calculus

### 2015-01-30 paper/implementation.tex
edits all over (syntax) and masking section
some work on stage one splitting and general syntax stuff
various edits and renames

We have implemented the semantics and splitting algorithm of \lang\ in Standard ML.
In addition to the features shown here, our implementation includes more base types, 
$n$-ary products and sums, if and let statements, and a basic elaboration step.

@@

We choose to investigate the latter approach.
Specifically, we add new staging features to \lang\ which allow us to indicate that a piece of code is {\em strictly} 
stage \bbone---{\em i.e.} that it contains no $\next$ blocks.
At the type level, this manifests as a constructor called $\monoType$.
So for instance, $\rmint \to \rmint$ is a stage \bbone\ function from one integer to another,
which might fail to terminate at stage \bbtwo.  
Alternatively $\monoType (\rmint \to \rmint)$ is a stage \bbone\ function from one integer to another
which {\em cannot} have any content at stage \bbtwo.
Naturally, the well-formedness of $\monoType A$ requires that $A$ contain no $\fut$s.

At the term level, we add the $\monoTerm$, $\lifttag$, \texttt{let mono} constructs to move information
in and out of $\monoType$.

\input{figures/mono.tex}

>>>>

\begin{concretesyntax}
We have implemented the semantics and splitting algorithm of \lang\ in Standard ML.
In addition to the features shown here, our implementation includes more base types, 
$n$-ary products and sums, if and let statements, and a basic elaboration step.

@@

We choose to investigate the latter approach.
Specifically, we add new staging features to \lang\ which allow us to indicate that a piece of code is {\em strictly} 
stage \bbone---{\em i.e.} that it contains no $\next$ blocks.
At the type level, this manifests as a constructor called $\curr$.
So for instance, $\rmint \to \rmint$ is a stage \bbone\ function from one integer to another,
which might fail to terminate at stage \bbtwo, 
but $\curr (\rmint \to \rmint)$ is a stage \bbone\ function from one integer to another
which {\em cannot} have any content at stage \bbtwo.
Naturally, the well-formedness of $\curr A$ requires that $A$ contain no $\fut$s.

With these new features in place, we can now define the \texttt{double}
function so that it splits reasonably.
\begin{lstlisting}
2`prev{
  1`let mono{`3`double`1`} = mono{`3`fn n : int => n+n`1`}
  in next{`2`hold {`1`lift(mono{`3`double 5`1`})`2`}-7`1`}`2`
}`
\end{lstlisting}
This now splits into:
\begin{lstlisting}
1`let double = (fn n : int => n+n)
in (double 5,())`

2`l => in (pi1 l)-7`
\end{lstlisting}
\end{concretesyntax}
